linguistic_analysis_C-D.Rmd : Linguistic quantity and complexity

### 2021-09-06 code/aa-flowers-pilot_analysis.Rmd
add prelim data tweaks and ms analyses

---
title: "flowers-pilot-analysis"
output:
  html_document: default
  pdf_document: default
---

@@

source(here("code/preprocessing/00_preprocessing_functions.R"))
source(here("code/theme_specs.R"))
#add pretty fonts
library(showtext)
font_add_google(name = "Assistant", family = "Assistant")
font_add_google(name = "Roboto", family = "Roboto")
showtext_auto()

@@

# Pre-process
Remove joined_data and uncomment the following block to rerun the preprocessing scripts
```{r process-data, message=FALSE, warning=FALSE}
data_list <- c("pilots", "collab_lang_20", "compet_lang_20", "nonlang_pilots", "compet_nonlang_15", "collab_nonlang_15")
process_all_data(data_list)
#join_all_data()
```

@@

Rename conditions for readability
```{r}
d.games <- d.games %>% mutate(conditionName = condition, 
                              condition = case_when(condition == "coopMulti" ~ "Cooperative",
                                         condition == "competCartel" ~ "Competitive",
                                         condition == "coopCartel" ~ "Old Cooperative",
                                         TRUE ~ condition))
d.games
```

@@

Game counts
```{r}
#pull out pilots
d.rounds.final %>% left_join(d.games %>% select(gameId, chatEnabled)) %>% group_by(condition, chatEnabled) %>% 
  summarize(n_games = n_distinct(gameId)) %>% arrange(condition, chatEnabled)
```

# Demographics
```{r demographics}
d.demographics
```

@@

@@

## Simple Distributions
### Individual Points
```{r, fig.height=3, fig.width=5}
d.by_player_metrics %>%
  group_by(languageCondition, condition, playerId) %>%
  summarize(playerPoints = sum(playerUtility)) %>%
  ggplot(aes(x = languageCondition, y = playerPoints, color = condition)) + 
  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + scale_color_bay() + theme_jmank() + theme(legend.position = "bottom")
```

@@

#Points per individual (per trial)

@@

```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = pointsAdjusted, color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Language Condition", 
       y = "Points Earned by Group (Max Adjusted)", 
       title = "Points by Group") + 
  ylim(0,1) +
  scale_color_bay() + theme_jmank() + theme(legend.position = "none")

@@

```
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = log(gameNumWords+1), color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition))+ 
  labs(y = "Log N. Words Exchanged")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "numwords_log_distirbution.png"))

@@

## Covariate
### Time x Points
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(y = gamePoints, x = gameLength/60, color = condition, shape = languageCondition)) + 
  geom_point(alpha = .5) +
  geom_smooth(method=lm)+ 
  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "time_words_scatter.png"))
```
# All Correlated
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = gameLength/60, y = gamePoints, shape = languageCondition, color = log(gameNumWords+1))) + 
  geom_point(alpha = .8) +
  geom_smooth(aes(group = condition), method=lm, color = "black")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Game length (in Minutes)", 
       y = "Points Earned by Group",
       color = "Log Num. Words",
       shape = "Language") + 
  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555")

@@

>>>>

---
title: "flowers-pilot-analysis"
output:
  html_document: default
  pdf_document: default
---
This notebook runs the preprocessing scripts, saves the data, and generates many summary analyses and plots

@@

source(here("code/preprocessing/00_preprocessing_functions.R"))
source(here("code/theme_specs.R"))
#add pretty fonts
library(showtext)
#font_add_google(name = "Assistant", family = "Assistant")
#font_add_google(name = "Roboto", family = "Roboto")
showtext_auto()

@@

# Pre-process
Remove joined_data and uncomment the following block to rerun the preprocessing scripts
```{r process-data, message=FALSE, warning=FALSE}
data_list <- c("pilots", "collab_lang_20", "compet_lang_20", "nonlang_pilots", "compet_nonlang_15", "collab_nonlang_15")
#process_all_data(data_list)
#join_all_data()
```

@@

Rename conditions for readability
```{r}
d.games <- d.games %>% mutate(conditionName = condition, 
                              condition = case_when(condition == "coopMulti" ~ "Shared Utilities",
                                         condition == "competCartel" ~ "Individual Utilities",
                                         condition == "coopCartel" ~ "Old Cooperative",
                                         TRUE ~ condition))
d.games
```

@@

Game counts
```{r}
#pull out pilots
game_completion %>% group_by(condition, chat, gameComplete) %>% 
  summarize(n_games = n_distinct(gameId)) %>% arrange(-gameComplete, -chat, condition)
```

# Demographics
```{r demographics}
d.demographics_clean <- d.demographics %>% filter(gameId %in% game_completion$gameId) %>%
  mutate(gender_code = case_when(gender %in% c("Female", "female", "f", "Woman","Femaile", "FEMALE", "Trans female", "Femal", "woman") ~ "Female",
                                gender %in% c("Male", "male", "MALE") ~ "Male",
                                is.na(gender) ~ "Unspecified",
                                gender %in% c("Non-Binary", "non-binary", "Non-binary")~ "Non-Binary", gender %in% c("agender") ~ "Agender", TRUE ~ "Other"))

d.demographics_clean %>% count(gender_code)

nrow(d.demographics_clean)
```

@@

## Word Counts
```{r}
d.raw_chat.final %>% filter(!is.na(text))

sum(d.by_game_metrics$gameNumWords)
```

@@

## Simple Distributions
#Points per individual (per trial) 
```{r, fig.height=3, fig.width=5}
#average points earned by participants per round
player_points_earned <- d.by_player_metrics %>%
  group_by(languageCondition, condition, playerId) %>%
  summarize(playerPoints = mean(playerUtility)) %>%
  ggplot(aes(x = languageCondition, y = playerPoints, color = condition)) + 
  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
  scale_color_bay() + theme_jmank() + 
  labs(y = "Average Player Points", x = "", color = "Utility Condition") +theme(legend.position = "bottom")

player_points_earned
```

@@

@@

```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = pointsAdjusted, color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Language Condition", 
       y = "Points Earned by Group (Max Adjusted)", 
       title = "Points by Group") + 
  #ylim(0,1) +
  scale_color_bay() + theme_jmank() + theme(legend.position = "none")

@@

@@

```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = log(gameNumWords+1), color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition))+ 
  labs(y = "Log N. Words Exchanged")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "numwords_log_distirbution.png"))

## Covariate
### Time x Points
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(y = gamePoints, x = gameLength/60, color = condition, shape = languageCondition)) + 
  geom_point(alpha = .5) +
  geom_smooth(method=lm)+ 
  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "time_words_scatter.png"))
```
# All Correlated
```{r, fig.height=3, fig.width=5}
all_correlated <- d.by_game_metrics %>%
  ggplot(aes(x = gameLength/60, y = gamePoints, shape = languageCondition, color = log(gameNumWords+1))) + 
  geom_point(alpha = .8) +
  geom_smooth(aes(group = condition), method=lm, formula=y~log(x),color = "black")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Game length (in Minutes)", 
       y = "Points Earned by Group",
       color = "Log N. Words",
       shape = "Language") + 
  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555")

all_correlated

@@

# MS Figs

```{r}

library(cowplot)
all_correlated
player_points_earned

plot_grid(player_points_earned,all_correlated , labels = c('A', 'B'), label_size = 12)

ggsave(here(fig_path, "combined_fig.png"), height = 3, width = 10, rel_)
```

### 2021-09-06 code/linguistic_analysis_C-D.Rmd
add nonlang conditions for comparison
add prelim data tweaks and ms analyses

#jpeg("~/Desktop/plots/mlu_block.jpg", width = 950, height = 950)
ggplot(d_flowers_utt_length_block, aes(x=blockNum, y=mlu,  color=condition))+
#  facet_grid( rows =vars(condition)) +
  geom_jitter(alpha=.05)+
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  stat_summary(fun.data = "mean_cl_boot")+
 # scale_y_continuous(limits = c(0,15))+
  labs(title="Mean Length of Utterance", y="MLU", x="Block number")+
  theme(legend.position="bottom") + 
  theme_jmank() + scale_color_bay() + theme(legend.position = "bottom")
#  theme(text = element_text(size = 30))
#dev.off() 
ggsave("../figs/mean_utt_length_block.png", height = 3, width = 3)

@@

### 3. Parts of speech
```{r pos}

text <- d_flowers$text
parsed <- spacy_parse(text,pos=TRUE) %>%
  select(doc_id, pos) %>%
  group_by(doc_id) %>%
  mutate(pos = paste(pos, collapse=",")) %>%
  distinct()

d_flowers_pos  <- cbind(d_flowers, parsed) %>%
  separate_rows(pos, convert = TRUE) %>%
  group_by(condition, blockNum, gameId, pos) %>%
  filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
  summarise(pos_count=n()) %>%
  left_join(d_flowers_utt_length_block) %>%
  mutate(prop_pos=pos_count/total_num_words)

#jpeg("~/Desktop/plots/pos_block.jpg", width = 950, height = 950)
ggplot(d_flowers_pos, 
       aes(x=blockNum, y=prop_pos, label=gameId,  color=condition)) + 
  geom_smooth(method = "lm", formula = y~x) + 
  geom_point(alpha=.1) +
  facet_grid( cols=vars(pos)) +
  xlab("Block number") + 
  ylab("Proportion of pos") + 
  theme_bw()#+
 # theme(text = element_text(size = 30))
#dev.off()

d_flowers_pos  <- cbind(d_flowers, parsed) %>%
  separate_rows(pos, convert = TRUE) %>%
  group_by(condition, trialNum, gameId, pos) %>%
  filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
  summarise(pos_count=n()) %>%
  left_join(d_flowers_utt_length_trial) %>%
  mutate(prop_pos=pos_count/total_num_words)

#jpeg("~/Desktop/plots/pos_trial.jpg", width = 950, height = 950)
ggplot(d_flowers_pos, 
       aes(x=trialNum, y=prop_pos, label=gameId,  color=condition)) + 
  geom_smooth(method = "lm", formula = y~x) + 
  geom_point(alpha=.1) +
  facet_grid( cols=vars(pos)) +
  xlab("Trial number") + 
  ylab("Proportion of pos") + 
  ylim(0,5)+
  theme_bw()#+
  #theme(text = element_text(size = 30))
#dev.off()

d_flowers_pos  <- cbind(d_flowers, parsed) %>%
  separate_rows(pos, convert = TRUE) %>%
  group_by(condition, repNum, gameId, pos) %>%
  filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
  summarise(pos_count=n()) %>%
  left_join(d_flowers_utt_length_rep) %>%
  mutate(prop_pos=pos_count/total_num_words)

#jpeg("~/Desktop/plots/pos_rep.jpg", width = 950, height = 950)
ggplot(d_flowers_pos, 
       aes(x=repNum, y=prop_pos, label=gameId, color=condition)) + 
  geom_smooth(method = "lm", formula = y~x) + 
  geom_point(alpha=.1) +
  facet_grid(cols=vars(pos)) +
  xlab("Rep number") + 
  ylab("Proportion of pos") +
  ylim(0,5)+
  theme_bw()#+
 # theme(text = element_text(size = 30))
#dev.off() 
```

@@

```{r minimum_edit_distance, include=TRUE}

>>>>

#jpeg("~/Desktop/plots/mlu_block.jpg", width = 950, height = 950)
ggplot(d_flowers_utt_length_block, aes(x=blockNum, y=mlu,  color=condition))+
#  facet_grid( rows =vars(condition)) +
  geom_jitter(alpha=.05)+
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  stat_summary(fun.data = "mean_cl_boot")+
 # scale_y_continuous(limits = c(0,15))+
  labs(title="Mean Length of Utterance", y="MLU", x="Block number")+
  theme(legend.position="bottom") + 
  theme_jmank() + scale_color_bay() + theme(legend.position = "bottom")
#  theme(text = element_text(size = 30))
#dev.off() 
#ggsave("../figs/mean_utt_length_block.png", height = 3, width = 3)

@@

### 3. Parts of speech
```{r pos, eval=FALSE, include=FALSE}

# text <- d_flowers$text
# parsed <- spacy_parse(text,pos=TRUE) %>%
#   select(doc_id, pos) %>%
#   group_by(doc_id) %>%
#   mutate(pos = paste(pos, collapse=",")) %>%
#   distinct()
#   
# d_flowers_pos  <- cbind(d_flowers, parsed) %>%
#   separate_rows(pos, convert = TRUE) %>%
#   group_by(condition, blockNum, gameId, pos) %>%
#   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
#   summarise(pos_count=n()) %>%
#   left_join(d_flowers_utt_length_block) %>%
#   mutate(prop_pos=pos_count/total_num_words)
#         
# #jpeg("~/Desktop/plots/pos_block.jpg", width = 950, height = 950)
# ggplot(d_flowers_pos, 
#        aes(x=blockNum, y=prop_pos, label=gameId,  color=condition)) + 
#   geom_smooth(method = "lm", formula = y~x) + 
#   geom_point(alpha=.1) +
#   facet_grid( cols=vars(pos)) +
#   xlab("Block number") + 
#   ylab("Proportion of pos") + 
#   theme_bw()#+
#  # theme(text = element_text(size = 30))
# #dev.off() 
# 
# d_flowers_pos  <- cbind(d_flowers, parsed) %>%
#   separate_rows(pos, convert = TRUE) %>%
#   group_by(condition, trialNum, gameId, pos) %>%
#   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
#   summarise(pos_count=n()) %>%
#   left_join(d_flowers_utt_length_trial) %>%
#   mutate(prop_pos=pos_count/total_num_words)
#         
# #jpeg("~/Desktop/plots/pos_trial.jpg", width = 950, height = 950)
# ggplot(d_flowers_pos, 
#        aes(x=trialNum, y=prop_pos, label=gameId,  color=condition)) + 
#   geom_smooth(method = "lm", formula = y~x) + 
#   geom_point(alpha=.1) +
#   facet_grid( cols=vars(pos)) +
#   xlab("Trial number") + 
#   ylab("Proportion of pos") + 
#   ylim(0,5)+
#   theme_bw()#+
#   #theme(text = element_text(size = 30))
# #dev.off() 
# 
# d_flowers_pos  <- cbind(d_flowers, parsed) %>%
#   separate_rows(pos, convert = TRUE) %>%
#   group_by(condition, repNum, gameId, pos) %>%
#   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
#   summarise(pos_count=n()) %>%
#   left_join(d_flowers_utt_length_rep) %>%
#   mutate(prop_pos=pos_count/total_num_words)
#         
# #jpeg("~/Desktop/plots/pos_rep.jpg", width = 950, height = 950)
# ggplot(d_flowers_pos, 
#        aes(x=repNum, y=prop_pos, label=gameId, color=condition)) + 
#   geom_smooth(method = "lm", formula = y~x) + 
#   geom_point(alpha=.1) +
#   facet_grid(cols=vars(pos)) +
#   xlab("Rep number") + 
#   ylab("Proportion of pos") +
#   ylim(0,5)+
#   theme_bw()#+
#  # theme(text = element_text(size = 30))
# #dev.off() 
```

@@
